How to obtain the redshift distribution from probabilistic redshift estimates

### 2017-08-01 research/paper/draft-editing.tex

A photo-$z$ PDF is then an \textit{interim posterior} probability distribution of the redshift of a galaxy given its observed photometric data and the interim prior: $p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})$.  We would like to learn about $p(z | \vec{\phi})$ from a catalog of photo-$z$ interim posteriors.

@@

We present the two attributes of a probabilistic graphical model (PGM) for the redshift distribution $n(z)$: the directed acyclic graph (DAG) and its mathematical interpretation in terms of Bayesian hierarchical inference.

\begin{figure}
	\begin{center}
		\includegraphics[width=0.5\textwidth]{pgm.png}
		\caption{This directed acyclic graph corresponds to a probabilistic graphical model for a Bayesian hierarchical inference of $p(\vec{\phi}|\{\vec{d}_{i}\})$.  In this graph, all random variables are shown in circles, with observed random variables shown in shaded circles.  Relationships between variables are indicated by arrows from parameters to the variables distributed according to functions of them.  The box indicates that there are a number of copies of the relationships between boxed parameters, each independent of all others.  The hyperparameters comprising $\vec{\phi}$, which define $n(z)$, are at the top.  Independently drawn from a function of the hyperparameters $\vec{\phi}$ are galaxy redshifts $\{z_{i}\}$ below.  The observed galaxy photometry $\{\vec{d}_{i}\}$, shown in shaded circles, is determined by the redshifts above.}
	\label{fig:pgm}
	\end{center}
\end{figure}

@@

\appendix{}
%showing stacking is always too broad

>>>>

A photo-$z$ PDF is then an \textit{interim posterior} probability distribution of the redshift of a galaxy given its observed photometric data and the interim prior: $p(z_{i} | \vec{d}_{i}, \vec{\phi}^{*})$.  
We present the two attributes of a probabilistic graphical model (PGM) for the redshift distribution $n(z)$: the directed acyclic graph (DAG) and its mathematical interpretation in terms of Bayesian hierarchical inference.

\begin{figure}
	\begin{center}We would like to learn about $p(z | \vec{\phi})$ from a catalog of photo-$z$ interim posteriors.

@@

		\includegraphics[width=0.5\textwidth]{fig/pgm.png}
		\caption{This directed acyclic graph corresponds to a probabilistic graphical model for a Bayesian hierarchical inference of $p(\vec{\phi}|\{\vec{d}_{i}\})$.  In this graph, all random variables are shown in circles, with observed random variables shown in shaded circles.  Relationships between variables are indicated by arrows from parameters to the variables distributed according to functions of them.  The box indicates that there are a number of copies of the relationships between boxed parameters, each independent of all others.  The hyperparameters comprising $\vec{\phi}$, which define $n(z)$, are at the top.  Independently drawn from a function of the hyperparameters $\vec{\phi}$ are galaxy redshifts $\{z_{i}\}$ below.  The observed galaxy photometry $\{\vec{d}_{i}\}$, shown in shaded circles, is determined by the redshifts above.}
	\label{fig:pgm}
	\end{center}
\end{figure}

@@

\appendix{Catalog Production}
\label{app:catalog}
%showing stacking is always too broad

Sec. \ref{sec:mockdata} outlines one way to make mock data, but it's not necessarily a logically consistent way to do it, a problem identified in trying to use it for the test cases in Sec. \ref{sec:likelihoods}.  If the goal is to emulate the traditional $z_{spec}$ vs. $z_{phot}$ scatterplots, we need to start from probability distributions in that space.  This means there are two parameters $z_{spec}$ and $z_{phot}$ drawn from a two-dimensional distribution $p(z_{spec}, z_{phot} | \textul{\sigma}, \vec{\phi})$, where $\textul{\sigma}$ contains parameters concerning the relationship between $z_{spec}$ and $z_{phot}$, and $\vec{\phi}$ contains hyperparameters concerning $n(z)$.  We know from basic probability that
\begin{align}
\label{eq:trivial}
p(z_{spec}, z_{phot} | \textul{\sigma}, \vec{\phi}) =& p(z_{phot} | z_{spec}, \textul{\sigma}, \vec{\phi})\ p(z_{sped} | \textul{\sigma}, \vec{\phi}).
\end{align}
Because $z_{spec}$ depends only on $\vec{\phi}$ and $z_{phot}$ is independent of $\vec{\phi}$, this becomes
\begin{align}
\label{eq:simpler}
p(z_{spec}, z_{phot} | \textul{\sigma}, \vec{\phi}) =& p(z_{phot} | z_{spec}, \textul{\sigma})\ p(z_{spec} | \vec{\phi}).
\end{align}
We know how to make reasonable choices for both terms on the righthand side of Eq. \ref{eq:simpler}.

In the $z_{spec}$ vs. $z_{phot}$ scatterplot, $p(z_{phot} | z_{spec}, \textul{\sigma}, \vec{\phi})$ would be represented by vertical slices, but we are interested in the horizontal slices that represent $p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi})$.  Now we must distinguish the true hyperparameters $\vec{\phi}_{true}$ from the interim hyperparameters $\vec{\phi}_{int}$.  If the horizontal slices carry information about $\vec{\phi}$, how can we get a catalog of $p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi}_{int})$ that's consistent with $p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi}_{true})$?  Somehow that information needs to be propagated through in order to recover it.

We achieve this goal by dividing out the true $n(z)$ and multiplying by the interim $n(z)$ as in
\begin{align}
\label{eq:transformation}
p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi}_{int}) =& p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi}_{true})\ \frac{p(z_{spec} | \vec{\phi}_{int})}{p(z_{spec} | \vec{\phi}_{true})},
\end{align}
since $p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi})$ is separable and the term $p(z_{spec} | z_{phot}, \textul{\sigma})$ changes independently of the $p(z_{spec} | \vec{\phi})$ term.  The information about $\vec{\phi}_{true}$ is retained by the elements of the catalog: only $(z_{spec}, z_{phot})$ pairs drawn from the true distribution will be represented in the catalog!

It appears as if this is the only fully self-consistent way to ensure that a draw from the posterior distribution $p(z_{spec} | z_{phot}, \textul{\sigma}, \vec{\phi})$ does in fact follow that distribution.  In the cases with fiducial posteriors $p(z_{spec} | z_{phot}, \sigma, \vec{\phi})$, the method of Sec. \ref{sec:mockdata} was equivalent to this, but Sec. \ref{sec:likelihoods} requires a more sophisticated approach.
